---
title: '04.2'
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Setup 

```{r}
"%&%" <- function(a,b) paste0(a,b)
library("data.table")
library("tidyverse")
serv.dir <- ""
base.dir <- serv.dir %&% "popgen/"
work.dir <- serv.dir %&%
  "popgen/02.2_admixture/afr_eas_eur_mais_amr-select-mcps1k/"
pop.file <- serv.dir%&% "shared/reference_datasets/" %&%
  "mais_information/reference-population-labels.txt"
```

```{r}
local.dir <- ""
```


```{r}
mais.ref.df <- fread(serv.dir%&% "shared/reference_datasets/" %&%
        "mais_information/mais-population-info_NJtree-regions.txt")
```


```{r}
ref.df <- readRDS(file= work.dir %&% "output_files/" %&%
          "ref.df.RDS")
```

Lookup MPCS district-specific samples 
```{r}
mcps.sub.df <- readRDS(file=work.dir %&% "output_files/mcps.sub.df.RDS")
```


```{r}
pop.df <- fread(pop.file)
pop.num <- purrr::map(mais.ref.df$mais.population,
                      function(s){
   dim(filter(pop.df,population==s))[1] 
}) %>% as.integer(.)
mais.ref.df$pop.num <- pop.num
mais.ref.df$Region <- factor(mais.ref.df$Region,
  levels=c("North","Northwest","Central","South",
           "Southeast"))
mais.show.vec <- (filter(mais.ref.df,
  pop.num>=10) %>% arrange(.,Region))$mais.population
```

African, East Asian, and European n>=10 populations 

```{r}
afr.sub.df <- filter(pop.df,region=="AFRICA")
afr.pop.df <- data.frame("Population"=unique(afr.sub.df$population))
afr.pop.df$pop.num <- purrr::map(unique(afr.sub.df$population),function(s){
  dim(filter(afr.sub.df,population==s))[1]
}) %>% as.integer(.)
afr.show.vec <- (filter(afr.pop.df,
  pop.num>=30))$Population
eas.sub.df <- filter(pop.df,region=="EAST_ASIA")
eas.pop.df <- data.frame("Population"=unique(eas.sub.df$population))
eas.pop.df$pop.num <- purrr::map(unique(eas.sub.df$population),function(s){
  dim(filter(eas.sub.df,population==s))[1]
}) %>% as.integer(.)
eas.show.vec <- (filter(eas.pop.df,
  pop.num>=20))$Population
eur.sub.df <- filter(pop.df,region=="EUROPE")
eur.pop.df <- data.frame("Population"=unique(eur.sub.df$population))
eur.pop.df$pop.num <- purrr::map(unique(eur.sub.df$population),function(s){
  dim(filter(eur.sub.df,population==s))[1]
}) %>% as.integer(.)
eur.show.vec <- (filter(eur.pop.df,
  pop.num>=30))$Population
```


## Functions

```{r}
get_top_groups <- function(mean.df,thresh=0.95){
  sort.df <- mean.df %>% sort(.,decreasing = TRUE)
  grp.names <- names(sort.df)
  grp.vec <- as.numeric(sort.df)
  cumsum = 0 
  keep.vec <- c()
  for (i in 1:length(grp.vec)){
    val <- grp.vec[i]
    if (cumsum<=thresh){
      keep.vec <- append(keep.vec,grp.names[i])
      cumsum <- cumsum + val 
    }
  }
  return(list(keep.vec,cumsum))
}

evaluate_thetas <- function(k,thresh=0.95){
  theta.ref.file <- work.dir%&%"output_files/ref."%&%k%&%".Q"
  theta.study.file <- work.dir%&%"output_files/mcps150k."%&%k%&%".Q"
  theta.ref.df <- fread(theta.ref.file,header=F)[,1:k]
  theta.study.df <- fread(theta.study.file,header=F)[,1:k]
  theta.df <- rbind(theta.ref.df,theta.study.df)
  names(theta.df) <- "G" %&% 1:k # "G" short for "Group"
  theta.df <- cbind(ref.df,theta.df)  
  # Find group averages by region in African, Europe, and Asia 
  afr.df <- filter(theta.df,region=="AFRICA")
  afr.sub.df <- dplyr::select(afr.df,contains("G"))
  afr.mean.df <- dplyr::summarise(afr.sub.df,
                        across(G1:("G"%&%k),mean))
  eas.df <- filter(theta.df,region=="EAST_ASIA")
  eas.sub.df <- dplyr::select(eas.df,contains("G"))
  eas.mean.df <- dplyr::summarise(eas.sub.df,
                        across(G1:("G"%&%k),mean))
  eur.df <- filter(theta.df,region=="EUROPE")
  eur.sub.df <- dplyr::select(eur.df,contains("G"))
  eur.mean.df <- dplyr::summarise(eur.sub.df,
                        across(G1:("G"%&%k),mean))
  afr.list <- get_top_groups(afr.mean.df,thresh)
  eas.list <- get_top_groups(eas.mean.df,thresh)
  eur.list <- get_top_groups(eur.mean.df,thresh)
  retain.vec <- c(afr.list[[1]],eas.list[[1]],eur.list[[1]])
  remain.vec <- ("G"%&%1:k)[!(("G"%&%1:k) %in% retain.vec)]
  afr.score <- dplyr::select(theta.df,one_of(afr.list[[1]])) %>% 
    rowSums(.)
  eas.score <- dplyr::select(theta.df,one_of(eas.list[[1]])) %>% 
    rowSums(.)
  eur.score <- dplyr::select(theta.df,one_of(eur.list[[1]])) %>% 
    rowSums(.)
  amr.score <- dplyr::select(theta.df,one_of(remain.vec)) %>% 
    rowSums(.)
  theta.df$afr.score <- afr.score
  theta.df$eas.score <- eas.score
  theta.df$eur.score <- eur.score
  theta.df$amr.score <- amr.score
  write.table(x=theta.df,file=work.dir%&%"output_files/"%&%
                "theta-k"%&%k%&%".txt",row.names=F,
              sep="\t",quote=F,col.names=T)
  # Order individuals by European score 
  ###samp.order <- arrange(theta.df,desc(eur.score))$sample
  # Order individuals by region, population, and ancestry scores 
  afr.order <- (filter(theta.df,region=="AFRICA") %>% 
    arrange(.,population,desc(afr.score)))$sample
  eas.order <- (filter(theta.df,region=="EAST_ASIA") %>% 
    arrange(.,population,desc(eas.score)))$sample 
  eur.order <- (filter(theta.df,region=="EUROPE") %>% 
    arrange(.,population,desc(eur.score)))$sample 
  amr.order.ref <- (filter(theta.df,region=="AMERICA",
                           population!="MCPS") %>% 
    arrange(.,population,desc(eur.score),desc(afr.score)))$sample
  amr.order.mcps <- (filter(theta.df,region=="AMERICA",
                           population=="MCPS") %>% 
    arrange(.,population,desc(eur.score),
            desc(afr.score)))$sample
  amr.order <- c(amr.order.ref,amr.order.mcps)
  samp.order <- c(afr.order,eas.order,
                          eur.order,amr.order)

  afr.order <- (filter(theta.df,region=="AFRICA") %>% 
    arrange(.,desc(afr.score)))$sample
  eas.order <- (filter(theta.df,region=="EAST_ASIA") %>% 
    arrange(.,desc(eas.score)))$sample 
  eur.order <- (filter(theta.df,region=="EUROPE") %>% 
    arrange(.,desc(eur.score)))$sample 
  amr.order <- (filter(theta.df,region=="AMERICA") %>% 
    arrange(.,desc(amr.score)))$sample 
  samp.order.revised <- c(afr.order,eas.order,
                          eur.order,amr.order)
  evaluate.thetas.list <- list(afr.list[[1]],eas.list[[1]],
                               eur.list[[1]],remain.vec,
                               samp.order,samp.order.revised)
  saveRDS(object = evaluate.thetas.list,
      file=work.dir%&%"output_files/evaluate.thetas.list"%&%
        "."%&%k%&%".thresh-"%&%thresh%&%".RDS")
  return(evaluate.thetas.list)
}
```

```{r}
library("RColorBrewer")
library("cowplot")
library("gridExtra")
col.func <- colorRampPalette(brewer.pal(11,
                             name="Spectral"))
```

```{r}

subgroup_plot <- function(k.sub.df,col.vec,hide.text=TRUE){
  vline.df <- c()
  count<-0
##  for (pop in levels(factor(k.sub.df$Population))){
  for (pop in levels(factor(k.sub.df$Population))){
    p.df <- filter(k.sub.df,Population==pop)
    reg <- unique(p.df$Region)
    ninds <- length(unique(p.df$IID))
    count<- count + ninds
    build.df <- data.frame("Region"=reg,"Population"=pop,
      "num.inds"=ninds,"index"=count,
      stringsAsFactors = F)
    vline.df <- rbind(vline.df,build.df)
  }
  pop.pos <- purrr::map(1:dim(vline.df)[1],function(i){
    if (i==1){
      pos <- round(0+(vline.df[i,]$index)/2,0)
    } else{
      pos <- round(((vline.df[i-1,]$index)+(vline.df[i,]$index))/2,0)
    }
  }) %>% as.integer(.)  
  vline.df$pop.pos <- pop.pos
  
  plt<- ggplot(data=k.sub.df,
                    aes(x=IID,y=Proportion)) + 
    geom_bar(stat="identity",aes(fill=K,col=K)) +  
    scale_fill_manual(values=col.vec,name="Group") + 
    scale_color_manual(values=col.vec,name="Group") + 
    scale_y_continuous(breaks=seq(0,1,0.2)) +
    theme(axis.text.x=element_blank(),
          axis.title.x = element_blank(),
          axis.ticks.x=element_blank(),
          panel.background = element_blank(),
          panel.grid.major = element_blank(),
          panel.grid.minor = element_blank(),
          strip.placement = "outside",
          legend.position ="none",
          strip.background = element_rect(fill="white"),
          strip.text = element_text(size=6)) + 
    ##ggtitle("K="%&%num.k) + 
    geom_vline(xintercept = vline.df$index,size=0.25,color="white") + 
    coord_cartesian(expand = F)# + 
  
  if (hide.text==TRUE){
    plt <- plt + theme(axis.text=element_blank(),
                       axis.title=element_blank(),
                       axis.ticks=element_blank())
  }
  
  vline.df$y.pos <- 1
  name.plt <- ggplot(data=vline.df,aes(x=pop.pos,y=y.pos)) + 
    geom_text(aes(label=Population),angle=90,size=2) + 
    theme(panel.border=element_blank(),panel.grid=element_blank(),
          axis.line=element_blank(),
          panel.background = element_blank(),
          axis.text=element_blank(),   
          axis.ticks=element_blank(),
          axis.title=element_blank(),
          #plot.margin = unit(c(0,-3.3,0,8), "mm"))
          plot.margin = unit(c(0,0,0,0), "mm"))

  combined.plot <- plot_grid(plt,name.plt,nrow=2,
                             rel_heights = c(2.5,1))
  return(combined.plot)
}

```

GGplot, ordered by region, population, and ancestry scores 
```{r}
ancestry_plot_unfaceted <- function(k.df,thresh=0.95,
                                    amr.col.vec=NULL,
                                    mcps=FALSE,split.district=FALSE,
                                    amr.group.order=NULL){
  num.k <- k.df$K %>% unique(.) %>% length(.)
  group.list <- evaluate_thetas(num.k,thresh)
  if (is.null(amr.group.order)){
      grp.order <- c(group.list[[1]],group.list[[2]],group.list[[3]],
                 group.list[[4]]) %>% unique(.)
  } else{
      grp.order <- c(group.list[[1]],group.list[[2]],group.list[[3]],
                 amr.group.order) %>% unique(.)
  }
  k.df$K <- factor(k.df$K,levels=grp.order)
  afr.num <- length(group.list[[1]])
  eas.num <- length(group.list[[2]])
  eur.num <- length(group.list[[3]])
  amr.num <- length(group.list[[4]])
  
  afr.colors <- rev(colorRampPalette(
    brewer.pal(9,name="YlOrRd"))(90)[20:30])
  afr.step.size <- round(length(afr.colors)/afr.num,0)
  afr.seq <- seq(1,length(afr.colors),afr.step.size)[1:afr.num]
  afr.colors <- afr.colors[afr.seq]
  
  eas.colors <- rev(colorRampPalette(
    brewer.pal(9,name="Reds"))(90)[60:70])
  eas.step.size <- round(length(eas.colors)/eas.num,0)
  eas.seq <- seq(1,length(eas.colors),eas.step.size)[1:eas.num]
  eas.colors <- eas.colors[eas.seq]
  
  eur.colors <- (colorRampPalette(
    brewer.pal(9,name="Blues"))(90)[30:70]) %>% rev(.)
  eur.step.size <- round(length(eur.colors)/eur.num,0)
  eur.seq <- seq(1,length(eur.colors),eur.step.size)[1:eur.num]
  eur.colors <- eur.colors[eur.seq]

  if (is.null(amr.col.vec)){
    amr.pal <- brewer.pal(11,name="PiYG")
    amr.colors <- (colorRampPalette(amr.pal)(100)) %>% rev(.)
    amr.step.size <- round(length(amr.colors)/amr.num,0)
    amr.seq <- seq(1,length(amr.colors),amr.step.size)[1:amr.num]
    if (is.na(amr.seq[length(amr.seq)])){
      amr.seq[length(amr.seq)] <- 100
    }
    amr.colors <- amr.colors[amr.seq]    
  } else{
    amr.colors <- amr.col.vec
  }

  
  # Need to add a hack for the case where Asian and African groups   #are the same
  if (all(group.list[[2]]==group.list[[1]])){
    cvec <- c(afr.colors,eur.colors,amr.colors)
  } else{
    cvec <- c(afr.colors,eas.colors,eur.colors,amr.colors)
  }

  samp.order <- group.list[[5]] #group.list[[6]]
  k.df$IID <- factor(k.df$IID,levels=samp.order)
  
  
  #Create Plots
  # Africa plot 
  k.sub.df <- filter(k.df,Region=="AFRICA",
                           Population%in%afr.show.vec)
  afr.plt <- subgroup_plot(k.sub.df,cvec,hide.text=FALSE)
  
  # European Plot 
  k.sub.df <- filter(k.df,Region=="EUROPE",Population%in%eur.show.vec)
  eur.plt <- subgroup_plot(k.sub.df,cvec)
  
  # East Asia Plot 
  ###eas.vec <- c("CHB")
  k.sub.df <- filter(k.df,Region=="EAST_ASIA",Population%in%eas.show.vec)
  eas.plt <- subgroup_plot(k.sub.df,cvec)
  
  # America Plot / MAIS
  k.sub.df <- filter(k.df,Region=="AMERICA",Population %in% 
                       mais.show.vec)
  k.sub.df$Population <- factor(k.sub.df$Population,levels=mais.show.vec)
  samp.order.mais <- arrange(k.sub.df,Population)$IID %>% unique(.)
  k.sub.df$IID <- factor(k.sub.df$IID,levels=samp.order.mais)
  mais.plt <- subgroup_plot(k.sub.df,cvec)
  
  ## MXL Admixed Plot 
  k.sub.df <- filter(k.df,Region=="AMERICA",Population=="MXL")
  mxl.plt <- subgroup_plot(k.sub.df,cvec,
                             hide.text=TRUE)
  combined.plt.top <- plot_grid(afr.plt,eas.plt,
            eur.plt,mais.plt,mxl.plt,
            nrow=1,rel_widths=c(5,3,3,15,3))
  
  if (mcps==TRUE){
    if (split.district==TRUE){
        ## MCPS: Coyoacan Admixed Plot 
        k.sub.df <- filter(k.df,Region=="AMERICA",
                           Population=="MCPS",IID %in%
            filter(mcps.sub.df,district=="Coyoacan")$sample)
        mcps.coy.plt <- subgroup_plot(k.sub.df,cvec,
                                   hide.text=FALSE)
        ## MCPS: Itztapalapa Admixed Plot 
        k.sub.df <- filter(k.df,Region=="AMERICA",
                           Population=="MCPS",IID %in%
            filter(mcps.sub.df,district=="Iztapalapa")$sample)
        mcps.izt.plt <- subgroup_plot(k.sub.df,cvec,
                                   hide.text=TRUE)
        combined.plt.bottom <- plot_grid(mcps.coy.plt,
                          mcps.izt.plt,nrow=1,rel_widths=c(5,6))
        combined.plt.full <- plot_grid(combined.plt.top,
                combined.plt.bottom,nrow=2)
        ##return(combined.plt.full)
        return(list(combined.plt.full,afr.colors,eas.colors,
                    eur.colors,group.list))
    } else{
        ## MCPS: all plot 
        k.sub.df <- filter(k.df,Region=="AMERICA",
                     Population=="MCPS",IID %in%
        filter(mcps.sub.df)$sample)
        mcps.plt <- subgroup_plot(k.sub.df,cvec,
                             hide.text=FALSE)
        combined.plt.bottom <- plot_grid(mcps.plt,nrow=1)
        combined.plt.full <- plot_grid(combined.plt.top,
                combined.plt.bottom,nrow=2)
        ##return(combined.plt.full)
        return(list(combined.plt.full,afr.colors,eas.colors,
                    eur.colors,group.list))
    }
  } else{
    ##return(combined.plt.top)
    return(list(combined.plt.top,afr.colors,eas.colors,
                eur.colors,group.list))
  }
}
```

### K=4 model 

```{r}
k4.df <- readRDS(file=work.dir%&%"output_files/k4.df.RDS")
```


```{r}
k4.plt.ref_only.list <- ancestry_plot_unfaceted(k4.df)
k4.plt.ref_only <- k4.plt.ref_only.list[[1]]
k4.plt.mcps.list <- ancestry_plot_unfaceted(k4.df,mcps=TRUE)
k4.plt.mcps <- k4.plt.mcps.list[[1]]
k4.plt.mcps_districts.list <- ancestry_plot_unfaceted(k4.df,
                            mcps=TRUE,split.district=TRUE)
k4.plt.mcps_districts <- k4.plt.mcps_districts.list[[1]]
```


Saving K=4 output 
```{r}
ggsave(x=k4.plt.ref_only,filename=work.dir%&%
"output_files/plots/k4_ref-only.plot.png",
height=3,width=16)
ggsave(x=k4.plt.mcps,filename=work.dir%&%
"output_files/plots/k4_mcps.plot.png",
height=6,width=16)
ggsave(x=k4.plt.mcps_districts,filename=work.dir%&%
"output_files/plots/k4_mcps-districts.plot.png",
height=6,width=16)

svg(file=local.dir%&%"output_files/plots/k4_ref-only.plot.svg",
    height =3,width = 16)
k4.plt.ref_only
dev.off()
```


### K=18 model 

```{r}
k18.df <- readRDS(file=work.dir%&%"output_files/k18.df.RDS")
```


K=18 plot 
```{r}
k=18
orange.vec <- brewer.pal(9,name="Oranges")
green.vec <- brewer.pal(9,name="YlGn")
purple.vec <- brewer.pal(9,name="Purples")
pink.vec <- brewer.pal(11,name="PiYG")
blue.vec <- brewer.pal(9,name="Blues")

theta.ref.file <- work.dir%&%"output_files/ref."%&%k%&%".Q"
theta.study.file <- work.dir%&%"output_files/mcps150k."%&%k%&%".Q"
theta.ref.df <- fread(theta.ref.file,header=F)[,1:k]
theta.study.df <- fread(theta.study.file,header=F)[,1:k]
theta.df <- rbind(theta.ref.df,theta.study.df)
names(theta.df) <- "G" %&% 1:k # "G" short for "Group"
theta.df <- cbind(ref.df,theta.df) 

group.list <- evaluate_thetas(18,thresh=0.95)
grp.order <- c(group.list[[1]],group.list[[2]],group.list[[3]],
                 group.list[[4]]) %>% unique(.)
group.list[[4]]


##Lookup up groups highest in Northern Mais
n.df <- filter(theta.df,population%in%filter(mais.ref.df,
                  Region=="North")$mais.population)
n.sub.df <- dplyr::select(n.df,contains("G"))
n.mean.df <- dplyr::summarise(n.df,
                      across(G1:("G"%&%k),mean)) 

##Lookup up groups highest in Northwest Mais
nw.df <- filter(theta.df,population%in%filter(mais.ref.df,
                  Region=="Northwest")$mais.population)
nw.sub.df <- dplyr::select(nw.df,contains("G"))
nw.mean.df <- dplyr::summarise(nw.df,
                      across(G1:("G"%&%k),mean)) 

##Lookup up groups highest in Central Mais
c.df <- filter(theta.df,population%in%filter(mais.ref.df,
                  Region=="Central")$mais.population)
c.sub.df <- dplyr::select(c.df,contains("G"))
c.mean.df <- dplyr::summarise(c.df,
                      across(G1:("G"%&%k),mean)) 

##Lookup up groups highest in Southern Mais
s.df <- filter(theta.df,population%in%filter(mais.ref.df,
                  Region=="South")$mais.population)
s.sub.df <- dplyr::select(s.df,contains("G"))
s.mean.df <- dplyr::summarise(s.df,
                      across(G1:("G"%&%k),mean))

##Lookup up groups highest in Southeaster Mais
se.df <- filter(theta.df,population%in%filter(mais.ref.df,
                  Region=="Southeast")$mais.population)
se.sub.df <- dplyr::select(se.df,contains("G"))
se.mean.df <- dplyr::summarise(se.df,
                      across(G1:("G"%&%k),mean))

(n.mean.df %>% sort(.,decreasing = TRUE))[c(1:3)]#G18
(nw.mean.df %>% sort(.,decreasing = TRUE))[c(1:3)]#G10
(c.mean.df %>% sort(.,decreasing = TRUE))[c(1:3)]#G9
(s.mean.df %>% sort(.,decreasing = TRUE))[c(1:3)]#G5
(se.mean.df %>% sort(.,decreasing = TRUE))[c(1:3)]#G6
## "G5"  "G6"  "G7"  "G9"  "G10" "G15" "G17" "G18"
fill.vec <- colorRampPalette(green.vec)(11)
manual.vec <- c("#FB9A99","#B3367AFF",fill.vec[3],
                "#FDBF6F","#FF7F00",fill.vec[7],fill.vec[10],orange.vec[9])
```

AMR-MAIS-North  "#E31A1C" / orange.vec[9]
AMR-MAIS-Northwest  "#FF7F00"
AMR-MAIS-Central  "#FDBF6F"
AMR-MAIS-South  "#FB9A99"
AMR-MAIS-Southeast  "#B3367AFF"

To make SE group easier to see in unfaceted plot, will manuall switch group 
order for Southern (G5) and SE (G6) group 
```{r}
amr.g.o <- c("G6","G5","G7","G9","G10","G15","G17","G18")
manual.vec <- c("#B3367AFF","#FB9A99",fill.vec[3],
                "#FDBF6F","#FF7F00",fill.vec[7],fill.vec[10],orange.vec[9])
```


```{r}
k18.plt.ref_only.list <- ancestry_plot_unfaceted(k18.df,
                            amr.col.vec=manual.vec,
                            amr.group.order=amr.g.o)
k18.plt.ref_only <- k18.plt.ref_only.list[[1]]
k18.plt.mcps.list <- ancestry_plot_unfaceted(k18.df,mcps=TRUE,
                    amr.col.vec=manual.vec,amr.group.order=amr.g.o)
k18.plt.mcps <- k18.plt.mcps.list[[1]]
k18.plt.mcps_districts.list <- ancestry_plot_unfaceted(k18.df,
                            mcps=TRUE,split.district=TRUE,
                      amr.col.vec=manual.vec,amr.group.order=amr.g.o)
k18.plt.mcps_districts <- k18.plt.mcps_districts.list[[1]]
```


```{r}
ggsave(x=k18.plt.ref_only,filename=work.dir%&%
"output_files/plots/k18_ref-only.plot.png",
height=3,width=16)
ggsave(x=k18.plt.mcps,filename=work.dir%&%
"output_files/plots/k18_mcps.plot.png",
height=6,width=16)
ggsave(x=k18.plt.mcps_districts,filename=work.dir%&%
"output_files/plots/k18_mcps-districts.plot.png",
height=6,width=16)

svg(file=local.dir%&%"output_files/plots/k18_ref-only.plot.svg",
    height =3,width = 16)
k18.plt.ref_only
dev.off()

```


## Faceted plots 

Functions
```{r}
region_plot <- function(sub.df,col.vec,hide.legend=FALSE){
  plt <- ggplot(data=sub.df,aes(x=IID,y=Proportion)) + 
    geom_bar(stat="identity",
             aes(fill=K,col=K)) + 
  scale_y_continuous(breaks=seq(0,1,0.1)) + 
  scale_fill_manual(values=col.vec) + 
  scale_color_manual(values=col.vec) + 
    facet_wrap(~Region,scales="free_x",
               strip.position="bottom",
               nrow=1) + 
    theme(axis.text.x=element_blank(),
          axis.title.x = element_blank(),
          axis.ticks.x=element_blank(),
          panel.background = element_blank(),
          panel.grid.major = element_blank(),
          panel.grid.minor = element_blank(),
          strip.placement = "outside",
          strip.background = element_rect(fill="white"),
          strip.text = element_text(size=6))
  if (hide.legend==TRUE){
    plt <- plt + theme(legend.position ="none")
  }
  return(plt)
}

pop_plot <- function(sub.df,col.vec,hide.text=TRUE,hide.legend=FALSE){
  plt <- ggplot(data=sub.df,aes(x=IID,y=Proportion)) + 
    geom_bar(stat="identity",
             aes(fill=K,col=K)) + 
  scale_y_continuous(breaks=seq(0,1,0.1)) + 
  scale_fill_manual(values=col.vec) + 
  scale_color_manual(values=col.vec) + 
    facet_wrap(~Population,scales="free_x",
               strip.position="bottom",
               nrow=1) + 
    theme(axis.text.x=element_blank(),
          axis.title.x = element_blank(),
          axis.ticks.x=element_blank(),
          panel.background = element_blank(),
          panel.grid.major = element_blank(),
          panel.grid.minor = element_blank(),
          strip.placement = "outside",
          strip.background = element_rect(fill="white"),
          strip.text = element_text(size=6))
  if (hide.text==TRUE){
    plt <- plt + theme(axis.text=element_blank(),
                       axis.title=element_blank(),
                       axis.ticks=element_blank())
  }
  if (hide.legend==TRUE){
    plt <- plt + theme(legend.position ="none")
  }
  return(plt)
}
```

### K=4
```{r}
vir.vec <- viridis::viridis(20)
cvec <- c(vir.vec[20],"red",vir.vec[8],vir.vec[15])
# Note: Refactoring to enable clear comparision with RFMix 3-way plot
k4.df$K <- factor(k4.df$K,levels=c("G3","G2","G4","G1"))
# Note: G3 corresponds to Europe in these results
lev.vec <- (filter(k4.df,K=="G4") %>% 
  arrange(desc(Proportion)))$IID %>% unique(.)
k4.df$IID <- factor(k4.df$IID,levels=lev.vec)

k4.plt.fac.ref <- region_plot(filter(k4.df,Population!="MCPS"),
                     col.vec=cvec,hide.legend=T)
k4.plt.fac.mcps <- pop_plot(filter(k4.df,Population=="MCPS"),
      col.vec=cvec,hide.text=F,hide.legend = F)
k4.plt.fac.full <- cowplot::plot_grid(k4.plt.fac.ref,k4.plt.fac.mcps,
      nrow=1,rel_widths = c(1,5))
k4.plt.fac.mcps <- cowplot::plot_grid(k4.plt.fac.mcps,nrow=1)
```

Save plots
```{r}
ggsave(plot=k4.plt.fac.mcps,filename=work.dir%&%
         "output_files/plots/admixture-plot-K4.png",
       height=2.5,width =12)
ggsave(plot=k4.plt.fac.full,filename=work.dir%&%
  "output_files/plots/admixture-plot-K4-with-references.png",
       height=2.5,width =16)
```

### K=18

AMR-MAIS-North  "#E31A1C" / orange.vec[9]
AMR-MAIS-Northwest  "#FF7F00"
AMR-MAIS-Central  "#FDBF6F"
AMR-MAIS-South  "#FB9A99"
AMR-MAIS-Southeast  "#B3367AFF"

To make SE group easier to see in unfaceted plot, will manually switch group 
order for Central (G9) before Southern (G5) before SE (G6) group 
```{r}
amr.g.o <- c("G15","G9","G5","G6","G7","G10","G17","G18")
manual.vec <- c(fill.vec[7],"#FDBF6F","#FB9A99","#B3367AFF",fill.vec[3],
                "#FF7F00",fill.vec[10],orange.vec[9])
```


```{r}
k18.list <- k18.plt.ref_only.list
lev.vec <- c(k18.list[[5]][[1]],k18.list[[5]][[2]],
             k18.list[[5]][[3]],amr.g.o)#k18.list[[5]][[4]]); manual tweak
color.vec <- c(k18.list[[2]],k18.list[[3]],k18.list[[4]],
               manual.vec)
k18.df$K <- factor(k18.df$K,levels=lev.vec)
samp.order <- rev(k18.list[[5]][[6]])
mais.pop.order <- arrange(mais.ref.df,Region)$mais.population
mais.order.df <- filter(ref.df,population%in%mais.pop.order)
mais.order.df$population <- factor(mais.order.df$population,
                                   levels=mais.pop.order)
mais.order <- arrange(mais.order.df,population)$sample
samp.order <- samp.order[!(samp.order%in%mais.order)]
samp.order <- c(samp.order,mais.order)

k18.df$IID <- factor(k18.df$IID,levels=samp.order)
  
k18.plt.fac.ref <- region_plot(filter(k18.df,Population!="MCPS"),
                     col.vec=color.vec,hide.legend=T)
k18.plt.fac.mcps <- pop_plot(filter(k18.df,Population=="MCPS"),
      col.vec=color.vec,hide.text=F,hide.legend = F) + 
  theme(legend.key.size=unit(0.7,"lines"))
k18.plt.fac.full <- cowplot::plot_grid(k18.plt.fac.ref,k18.plt.fac.mcps,
      nrow=1,rel_widths = c(1,5))
k18.plt.fac.mcps <- cowplot::plot_grid(k18.plt.fac.mcps,nrow=1)
```


```{r}
ggsave(plot=k18.plt.fac.mcps,filename=work.dir%&%
         "output_files/plots/admixture-plot-K18.png",
       height=3,width =12)
ggsave(plot=k18.plt.fac.full,filename=work.dir%&%
  "output_files/plots/admixture-plot-K18-with-references.png",
       height=3,width =16)
```