---
title: '01.1_prepare-input-files.Rmd'
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Setup 

```{r}
"%&%" <- function(a, b) paste0(a, b)
library("data.table")
library("tidyverse")
serv.dir <- ""
work.dir.serv <- serv.dir %&% 
  "popgen/04_rfmix/public_references_only/"
input.dir <- work.dir.serv %&% "input_files/"
output.dir <- work.dir.serv %&% "output_files/"
```

First need to extract sample ids from a vcf file 
```{bash}
cd popgen/03_phasing/output_files
module load BCFtools/1.10.2-GCC-8.3.0
bcftools query -l merged_mcps-hgdp-1kg_chr22_phased.vcf > ../../04_rfmix/public_references_only/input_files/sample-ids.txt
cd popgen/04_rfmix
```

```{r}
samp.df <- fread(input.dir %&% "sample-ids.txt", header=F)
names(samp.df)[1] <- "VCF.ID"
pb <- txtProgressBar(min=0, max=dim(samp.df)[1], style=3)
samp.df$ID <- purrr::map(1:dim(samp.df)[1], function(i){
  setTxtProgressBar(pb, i)
  name.string <- samp.df$VCF.ID[i]
  if (grepl("DUP", name.string) & !(grepl("FAM", name.string))){
    strsplit(name.string, split="-DUP-")[[1]][2]
  } else if(grepl("DUP", name.string) & grepl("FAM", name.string)){
    s.vec <- strsplit(name.string, split="-")[[1]]
    paste0(s.vec[2:length(s.vec)], collapse="-")
  } else{
    strsplit(name.string, split="-")[[1]][2]
  }
}) %>% as.character(.)
```

Merge with source population information (from PCA analysis) 

```{r}
pca.dir <- work.dir.serv <- serv.dir %&% 
  "popgen/01_pca/"
pca.df <- fread(pca.dir %&% "public/bigsnpr_output/" %&% 
  "pc_projections_unrelateds_maf01_pop-info.txt")
sub.df <- dplyr::select(pca.df, one_of("sample.ID", "population", "region"))
names(sub.df)[1] <- "ID"

full.df <- dplyr::inner_join(samp.df, sub.df, by="ID")

vec1 <- samp.df$ID[!(samp.df$ID %in% sub.df$ID)]
vec2 <- sub.df$ID[!(sub.df$ID %in% samp.df$ID)]
```

All samples accounted for between datasets 


## Create input sample files for RFMix 

Need cohort sample file

```{r}
mcps.vec <- full.df$VCF.ID[grepl("MCPS", full.df$VCF.ID)]
write.table(x=mcps.vec, file=input.dir %&%"mcps-samples.txt", sep="\t", 
            quote=F, row.names=F, col.names=F)
```

Need to select reference samples to keep,  and include in map file,  considering 
three-way admixture of ancestry from Africa, America, and Europe 

Accessing inferred ancestral population reference samples from TeraStructure

```{r}
tera.dir <- serv.dir %&% 
  "popgen/02_terastructure/freeze_145k/" 
afr.df <- fread(tera.dir %&%
  "output_files/teraK12.inferred-african-references.txt",header=F)
eur.df <- fread(tera.dir %&%
  "output_files/teraK12.inferred-european-references.txt",header=F)
amr.df <- fread(tera.dir %&%
  "output_files/teraK12.inferred-american-references.txt",header=F)
refsamp.df <- rbind(afr.df, eur.df, amr.df)
```

```{r}
#table(full.df$region)
ref.df <- filter(full.df, region %in% c("AFRICA", "EUROPE", "AMERICA"), 
                 population!="MCPS", ID%in%refsamp.df$V1)
write.table(x=ref.df$VCF.ID, file=input.dir %&%"ref-samples.txt", sep="\t", 
            quote=F, row.names=F, col.names=F)
write.table(x=dplyr::select(ref.df, one_of("VCF.ID", "region")), 
            file=input.dir %&%"ref-map.txt", sep="\t", 
            quote=F, row.names=F, col.names=F)
```


