---
title: '03.2'
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Setup 

```{r}
"%&%" <- function(a,b) paste0(a,b)
library("data.table")
library("tidyverse")
serv.dir <- "./"
work.dir.serv <- serv.dir %&%
  "popgen/02.1_terastructure/mais_amr_afr_eur_eas/"
tera.dir.pre1 <- "n3274-k"
tera.dir.pre2 <-"-l226827-select_k_"
tera.dir.suf <- "_rep"
```

https://github.com/StoreyLab/terastructure/wiki/Advanced-Usage

"Choosing K
We run the algorithm over a range of K with multiple reps for each value of K. Then, we extract the final validation 
likelihood for each run. Averaging over all reps, we choose the value of K where the validation likelihood plateaus."


```{r}
k.vec <- 4:20
```

```{r}
build_validation_df <- function(k.vec){
  out.df <- c()
  for (k in k.vec){
    val.vec <- c() 
    for (i in 1:3){
    df <- fread(work.dir.serv %&% "output_files/" %&% tera.dir.pre1 %&% 
                  k %&% tera.dir.pre2 %&% 
                  k %&% tera.dir.suf %&% i %&% "/validation.txt")
    final.lkhd <- df$V3[length(df$V3)]
    val.vec <- append(val.vec,final.lkhd)
    }
    build.df <- data.frame("K"=k,"avg.likelihood"=mean(val.vec),
                           "sd.likelihood"=sd(val.vec),
                           stringsAsFactors = F)
    out.df <- rbind(out.df,build.df)
  }
  return(out.df)
}
```

```{r}
val.df <- build_validation_df(k.vec)
```

```{r}
plt <- ggplot(data=val.df,
              aes(x=K,y=avg.likelihood)) + 
  geom_errorbar(aes(ymin=avg.likelihood-sd.likelihood,
                    ymax=avg.likelihood+sd.likelihood)) + 
  geom_line(col="black",size=2) + 
  scale_x_continuous(breaks=k.vec) + 
  theme_classic()
```

```{r}
library(gridExtra)
save.plt <- grid.arrange(plt,nrow=1)
ggsave(plot=save.plt,filename = work.dir.serv %&% "output_files/select-k.png",
       height=4,width=6)
```


After reviewing the average validation likelihood, it seems that K=18 would be the optimal number. 
Note that k=11, 15, and 18 all corresponded to peaks. 

