---
title: "04.6"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Setup

```{r}
"%&%" <- function(a,b) paste0(a,b)
library("data.table")
library("tidyverse")
library("bigsnpr")
library("RColorBrewer")
library("rgl")
serv.dir <- "./"
work.dir0 <- serv.dir %&% "popgen/01_pca/" %&%
  "public_mais/"
work.dir <- work.dir0 %&% "04_mais_rsq90/"
file.dir <- work.dir %&% "bigsnpr_output/"
```

## Look up model information from SVD objects

```{r}
svd.obj1 <- readRDS(file.dir %&% "svd_unrelateds_maf01.rds")
(svd.obj1$u %>% dim(.))[1] # 1306 subjects
(svd.obj1$v %>% dim(.))[1] # 78718 SNPs
pve.df1 <- data.frame("PC"="PC"%&%1:20,
                      "PVE"=((svd.obj1$d^2)/sum(svd.obj1$d^2))*100,
                      stringsAsFactors = F)
pve.df1$PVE <- pve.df1$PVE %>% prettyNum(digits=3)
rm(svd.obj1)

```

## Look up number of samples from each source used in the models and projections

```{r}
plink.dir <- work.dir0 %&% "merged_mcps/"
plink.pre <- "merged_reference_rsq90.merge.mcps.autosomes"
bed.file <- plink.dir %&% plink.pre %&% ".bed"
rel <- fread(work.dir0 %&%
  "04_mais_rsq90/bigsnpr_output/bigsnpr_relationships.txt")
obj.bed <- bed(bed.file)
```


```{r}
pc.df1 <- fread(file.dir%&%
              "pc_projections_unrelateds_maf01_pop-info.txt")
```

## Create PC Plots color-coded by population/ancestry


# Generate the PC scatter plots 

```{r}
col.vir <- viridis::viridis(20)
cvec <- brewer.pal(12,name="Paired")
pc.df1$plot.ID1 <- factor(pc.df1$plot.ID1,
 levels=as.character(unique(arrange(pc.df1,region,sample.type)$plot.ID1)))
reg.vec <- c("AMERICA","AFRICA","EUROPE")
```

rgl functions
```{r}
rgl_init <- function(new.device = FALSE, bg = "white", width = 640) { 
  if( new.device | rgl.cur() == 0 ) {
    rgl.open()
    par3d(windowRect = 50 + c( 0, 0, width, width ) )
    rgl.bg(color = bg )
  }
  rgl.clear(type = c("shapes", "bboxdeco"))
  rgl.viewpoint(theta = 15, phi = 20, zoom = 0.7)
}

rgl_add_axes <- function(x, y, z, axis.col = "grey",
                         xlab = "", ylab="", zlab="", show.plane = TRUE, 
                         show.bbox = FALSE, bbox.col = c("#333377","black"))
{ 
  
  lim <- function(x){c(-max(abs(x)), max(abs(x))) * 1.1}
  # Add axes
  xlim <- lim(x); ylim <- lim(y); zlim <- lim(z)
  rgl.lines(xlim, c(0, 0), c(0, 0), color = axis.col)
  rgl.lines(c(0, 0), ylim, c(0, 0), color = axis.col)
  rgl.lines(c(0, 0), c(0, 0), zlim, color = axis.col)
  
  # Add a point at the end of each axes to specify the direction
  axes <- rbind(c(xlim[2], 0, 0), c(0, ylim[2], 0), 
                c(0, 0, zlim[2]))
  rgl.points(axes, color = axis.col, size = 3)
  
  # Add axis labels
  rgl.texts(axes, text = c(xlab, ylab, zlab), color = axis.col,
            adj = c(0.5, -0.8), size = 2)
  

  # Add bounding box decoration
  if(show.bbox){
    rgl.bbox(color=c(bbox.col[1],bbox.col[2]), alpha = 0.5, 
             emission=bbox.col[1], specular=bbox.col[1], shininess=5, 
             xlen = 3, ylen = 3, zlen = 3) 
  }
}

get_colors <- function(groups, group.col = palette()){
  groups <- as.factor(groups)
  ngrps <- length(levels(groups))
  if(ngrps > length(group.col)) 
    group.col <- rep(group.col, ngrps)
  color <- group.col[as.numeric(groups)]
  names(color) <- as.vector(groups)
  return(color)
}
```




## 3D PCA Plots

### Plot 1: MAIS (all) + YRI + IBS + MCPS (5K random)
```{r}
movie.name <- "four-colors_mais_yri_ibs_mcps5K_fps15"
mcps.df <- filter(pc.df1,population=="MCPS")
# subset for plotting purposes
mcps.sub.df <- mcps.df[sample(1:dim(mcps.df)[1],size=5000,replace=F),]
reference.df1 <- filter(pc.df1,
  plot.ID2 %in% c("KG_AFR_YRI","KG_EUR_IBS"))
reference.df2 <- filter(pc.df1,
  plot.ID1 %in% c("MAIS_AMR"))
reference.df2$plot.ID2 <- "MAIS_AMR"
reference.df <- rbind(reference.df1,reference.df2)
plot.df <- rbind(mcps.sub.df,reference.df)
x <- dplyr::select(plot.df,one_of(pcx)) %>% unlist(.) %>% 
  as.numeric(.)
y <- dplyr::select(plot.df,one_of(pcy)) %>% unlist(.) %>% 
  as.numeric(.)
z <- dplyr::select(plot.df,one_of(pcz)) %>% unlist(.) %>% 
  as.numeric(.)
group.col.vec <- c(col.vir[20],col.vir[10],                     
                  col.vir[17],"gray25")
color.vec <- get_colors(plot.df$plot.ID2,group.col=group.col.vec)
rgl_init()
rgl.spheres(x, y, z, r = 2, color = color.vec) 
rgl_add_axes(x, y, z,xlab=pcx,ylab=pcy,zlab=pcz,
             show.bbox = FALSE,show.plane=FALSE,axis.col="gray40")
aspect3d(1,1,1)
#spin3d(axis = c(0,1,1))
movie3d(spin3d(axis = c(0,1,0)), duration = 12,
      dir = file.dir,movie=movie.name,fps=15)
```


### Plot 2a: MAIS (non-admixed) + YRI + IBS + zero MCPS 

Look up AMR MAIS samples with >=0.99 AMR ancestry
```{r}
tera.dir <- serv.dir %&% "popgen/02.1_terastructure/" %&% 
  "mais_analyses/imputed-topmed/mais_amr_afr_eur_eas_mcps10k/output_files/"
amr99.file <- tera.dir %&% "samples-k17-p99-amr.txt"
amr99.df <- fread(amr99.file,header=F)
```


```{r}
pcx<-"PC1";pcy<-"PC2";pcz<-"PC3"
movie.name <- "three-colors_mais660_yri_ibs_mcps0_fps15"
reference.df1 <- filter(pc.df1,
  plot.ID2 %in% c("KG_AFR_YRI","KG_EUR_IBS"))
reference.df2 <- filter(pc.df1,
  plot.ID1 %in% c("MAIS_AMR"))
reference.df2$plot.ID2 <- "MAIS_AMR"
reference.df2 <- filter(reference.df2,sample.ID %in% amr99.df$V1)
reference.df <- rbind(reference.df1,reference.df2)
plot.df <- rbind(reference.df) # mcps.sub.df,
x <- dplyr::select(plot.df,one_of(pcx)) %>% unlist(.) %>% 
  as.numeric(.)
y <- dplyr::select(plot.df,one_of(pcy)) %>% unlist(.) %>% 
  as.numeric(.)
z <- dplyr::select(plot.df,one_of(pcz)) %>% unlist(.) %>% 
  as.numeric(.)
group.col.vec <- c(col.vir[20],col.vir[10],                     
                  col.vir[17]) # ,"gray25"
color.vec <- get_colors(plot.df$plot.ID2,group.col=group.col.vec)
rgl_init()
rgl.spheres(x, y, z, r = 2, color = color.vec) 
rgl_add_axes(x, y, z,xlab=pcx,ylab=pcy,zlab=pcz,
             show.bbox = FALSE,show.plane=FALSE,axis.col="gray40")
aspect3d(1,1,1)
#spin3d(axis = c(0,1,1))
movie3d(spin3d(axis = c(0,1,0)), duration = 12,
      dir = file.dir,movie=movie.name,fps=15)
```

### Plot 2b: MAIS (non-admixed) + YRI + IBS + 10000 MCPS 


```{r}
pcx<-"PC1";pcy<-"PC2";pcz<-"PC3"
movie.name <- "four-colors_mais660_yri_ibs_mcps10K_fps15"
mcps.df <- filter(pc.df1,population=="MCPS")
# subset for plotting purposes
mcps.sub.df <- mcps.df[sample(1:dim(mcps.df)[1],size=10000,replace=F),]
reference.df1 <- filter(pc.df1,
  plot.ID2 %in% c("KG_AFR_YRI","KG_EUR_IBS"))
reference.df2 <- filter(pc.df1,
  plot.ID1 %in% c("MAIS_AMR"))
reference.df2$plot.ID2 <- "MAIS_AMR"
reference.df2 <- filter(reference.df2,sample.ID %in% amr99.df$V1)
reference.df <- rbind(reference.df1,reference.df2)
plot.df <- rbind(mcps.sub.df,reference.df) # 
x <- dplyr::select(plot.df,one_of(pcx)) %>% unlist(.) %>% 
  as.numeric(.)
y <- dplyr::select(plot.df,one_of(pcy)) %>% unlist(.) %>% 
  as.numeric(.)
z <- dplyr::select(plot.df,one_of(pcz)) %>% unlist(.) %>% 
  as.numeric(.)
group.col.vec <- c(col.vir[20],col.vir[10],                     
                  col.vir[17],"gray25")
color.vec <- get_colors(plot.df$plot.ID2,group.col=group.col.vec)
rgl_init()
rgl.spheres(x, y, z, r = 2, color = color.vec) 
rgl_add_axes(x, y, z,xlab=pcx,ylab=pcy,zlab=pcz,
             show.bbox = FALSE,show.plane=FALSE,axis.col="gray40")
aspect3d(1,1,1)
#spin3d(axis = c(0,1,1))
movie3d(spin3d(axis = c(0,1,0)), duration = 12,
      dir = file.dir,movie=movie.name,fps=15)
```
