---
title: "04.5"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Setup

```{r}
"%&%" <- function(a,b) paste0(a,b)
library("data.table")
library("tidyverse")
library("bigsnpr")
library("RColorBrewer")
library("gridExtra")
library("cowplot")
serv.dir <- "/Users/jasont/science/servers/FUSE1/"
work.dir0 <- serv.dir %&% "users/bjk420/projects/popgen/01_pca/" %&%
  "global_inmegen/imputed-topmed/"
work.dir <- work.dir0 %&% "04_mais_rsq90/mais_yri_ibs/"
file.dir <- work.dir %&% "bigsnpr_output/"
pc.df <- fread(file.dir %&% "pc_projections_unrelateds_maf01.txt")
```

## Append population labels to PC projection output

```{r}
onekg.df <- fread(serv.dir %&% "shared/reference_datasets/1000_genomes/"
 %&% "integrated_call_samples_v3.20130502.ALL.panel",
                  header=F) %>% dplyr::select(one_of("V1","V2","V3"))
hgdp.df <- fread(serv.dir %&% "shared/reference_datasets/" %&%
                   "hgdp_sanger/hgdp_wgs.20190516.metadata.txt",
                 header=T) %>%
  dplyr::select(one_of("sample","population","region"))
```

HGDP has 54 populations and seven regions
[1] "CENTRAL_SOUTH_ASIA" "AFRICA"             "OCEANIA"            "EUROPE"     
[5] "MIDDLE_EAST"        "AMERICA"            "EAST_ASIA"  

1KG has 26 populations and five superpopulations
[1] "EUR" "EAS" "AMR" "SAS" "AFR"

```{r}
onekg.df$V3 <- purrr::map(onekg.df$V3,function(s){
  ifelse(s=="EUR","EUROPE",s)
}) %>% as.character(.)
onekg.df$V3 <- purrr::map(onekg.df$V3,function(s){
  ifelse(s=="EAS","EAST_ASIA",s)
}) %>% as.character(.)
onekg.df$V3 <- purrr::map(onekg.df$V3,function(s){
  ifelse(s=="AMR","AMERICA",s)
}) %>% as.character(.)
onekg.df$V3 <- purrr::map(onekg.df$V3,function(s){
  ifelse(s=="SAS","CENTRAL_SOUTH_ASIA",s)
}) %>% as.character(.)
onekg.df$V3 <- purrr::map(onekg.df$V3,function(s){
  ifelse(s=="AFR","AFRICA",s)
}) %>% as.character(.)
names(onekg.df) <- c("sample","population","region")
```

MAIS datasets 

```{r}
mais.dir <- serv.dir %&% "projects/mcps/data/MAIS_array_data/"
affy.df <- fread(mais.dir %&% "MAIS.affy6.0.2021.ind",header=F) %>% 
  dplyr::select(.,one_of("V1","V3"))
omni.df <- fread(mais.dir %&% "MAIS.omni2.5.2021.ind",header=F) %>% 
  dplyr::select(.,one_of("V1","V3"))
names(affy.df) <- c("sample","population")
names(omni.df) <- c("sample","population")
affy.df$sample[!(affy.df$sample %in% pc.df$sample.ID)]
omni.df$sample[!(omni.df$sample %in% pc.df$sample.ID)]
omni.df$sample <- purrr::map(omni.df$sample,function(s){
  strsplit(x=s,split=":")[[1]][1]
}) %>% as.character(.)
omni.df$sample[!(omni.df$sample %in% pc.df$sample.ID)]
mais.df <- rbind(affy.df,omni.df)
mais.df$region <- "AMERICA"
```

```{r}
ref.df <- rbind(hgdp.df,onekg.df,mais.df)
```

```{r}
add_pop_info <- function(pc.df){
  pb <- txtProgressBar(min=0,max=dim(pc.df)[1],style=3)
  out.df <- c()
  for (i in 1:dim(pc.df)[1]){
    setTxtProgressBar(pb,i)
    row.df <- pc.df[i,]
    if (row.df$sample.ID %in% ref.df$sample){
      sub.df <- filter(ref.df,sample==row.df$sample.ID)
      names(sub.df)[1] <- "sample.ID"
      build.df <- dplyr::inner_join(sub.df,row.df,by="sample.ID")
    } else{
      sub.df <- data.frame("sample.ID"=row.df$sample.ID,
                           "population"="MCPS","region"="AMERICA",
                           stringsAsFactors = F)
      build.df <- dplyr::inner_join(sub.df,row.df,by="sample.ID")
    }
    out.df <- rbind(out.df,build.df)
  }
  return(out.df)
}


```

```{r}
pc.df1 <- add_pop_info(pc.df)
```

## Look up model information from SVD objects

```{r}
svd.obj1 <- readRDS(file.dir %&% "svd_unrelateds_maf01.rds")
(svd.obj1$u %>% dim(.))[1] # 1306 subjects
(svd.obj1$v %>% dim(.))[1] # 78718 SNPs
pve.df1 <- data.frame("PC"="PC"%&%1:20,
                      "PVE"=((svd.obj1$d^2)/sum(svd.obj1$d^2))*100,
                      stringsAsFactors = F)
pve.df1$PVE <- pve.df1$PVE %>% prettyNum(digits=3)
rm(svd.obj1)

```

## Look up number of samples from each source used in the models and projections

```{r}
plink.dir <- work.dir0 %&% "merged_mcps/"
plink.pre <- "merged_reference_rsq90.merge.mcps.autosomes"
bed.file <- plink.dir %&% plink.pre %&% ".bed"
rel <- fread(work.dir0 %&%
  "04_mais_rsq90/bigsnpr_output/bigsnpr_relationships.txt")
obj.bed <- bed(bed.file)
```

Analysis 1: HGDP, 1KG, MAIS, and MCPS samples used in PCA, projected to remaining MCPS

Model sample counts 
```{r}
keep.indices <- readRDS(file.dir %&% "keep.indices.RDS")
ind.row <- keep.indices#[S < 1.0] # No outlier thresholding is necessary here
subject.vec <- obj.bed$fam$sample.ID[keep.indices]
subject.df <- as.data.frame(subject.vec) # from 04.4 script
df1 <- subject.df %>% 
  filter(.,(grepl(x=subject.vec,pattern="HGDP"))) # 
df2 <- subject.df %>% 
  filter(.,(grepl(x=subject.vec,pattern="MCPS"))) #  
df3 <- subject.df %>% filter(.,!(grepl(x=subject.vec,pattern="HGDP"))) %>% 
  filter(.,grepl(x=subject.vec,pattern="HG")) #  
df4 <- subject.df %>% filter(.,!(grepl(x=subject.vec,pattern="HGDP"))) %>% 
  filter(.,grepl(x=subject.vec,pattern="NA")) # 
remove.vec <- rbind(df1,df2,df3,df4)$subject.vec
mais.df <- filter(subject.df,!(subject.vec%in%remove.vec)) # 591
samples.hgdp <- df1$subject.vec
samples.1kg.hg <- df3$subject.vec
samples.1kg.na <- df4$subject.vec
samples.1kg <- c(samples.1kg.hg,samples.1kg.na)
samples.mcps <- df2$subject.vec
samples.mais <- mais.df$subject.vec
samples.hgdp %>% length(.) # 0 HGDP samples
samples.1kg %>% length(.) # 215 1KG samples
samples.mcps %>% length(.) # 500 MCPS samples
```

Projected sample counts
```{r}
projected.count <- dim(pc.df1)[1]-length(subject.vec) # 141354
projected.samples <- pc.df1$sample.ID[!(pc.df1$sample.ID %in% subject.vec)]
projected.samples %>% length(.) # 141354 samples
proj.df <- as.data.frame(projected.samples)
names(proj.df) <- c("subject.vec")
pdf1 <- proj.df %>% 
  filter(.,(grepl(x=subject.vec,pattern="HGDP"))) # 
pdf2 <- proj.df %>% 
  filter(.,(grepl(x=subject.vec,pattern="MCPS"))) #  
pdf3 <- proj.df %>% filter(.,!(grepl(x=subject.vec,pattern="HGDP"))) %>% 
  filter(.,grepl(x=subject.vec,pattern="HG")) #  
pdf4 <- proj.df %>% filter(.,!(grepl(x=subject.vec,pattern="HGDP"))) %>% 
  filter(.,grepl(x=subject.vec,pattern="NA")) # 
remove.vec <- rbind(pdf1,pdf2,pdf3,pdf4)$subject.vec
mais.pdf <- filter(proj.df,!(subject.vec%in%remove.vec)) # 
projected.hgdp <- pdf1$subject.vec
projected.1kg.hg <- pdf3$subject.vec
projected.1kg.na <- pdf4$subject.vec
projected.1kg <- c(projected.1kg.hg,projected.1kg.na)
projected.mcps <- pdf2$subject.vec
projected.mais <- mais.pdf$subject.vec
projected.hgdp %>% length(.) # 929 HGDP samples
projected.1kg %>% length(.) # 2289 1KG samples
projected.mcps %>% length(.) # 138011 MCPS samples
projected.mais %>% length(.) # 125 MAIS samples

pb <- txtProgressBar(min=0,max=dim(pc.df1)[1],style=3)
pc.df1$sample.type <- purrr::map(1:dim(pc.df1)[1],function(i){
  setTxtProgressBar(pb,i)
  s <- pc.df1$sample.ID[i]
  ifelse(s%in%subject.vec,"model","projected")
}) %>% as.character(.)

```


```{r}
write.table(x=pc.df1,file=file.dir%&%
              "pc_projections_unrelateds_maf01_pop-info.txt",
            quote=F,sep="\t",row.names=F)
```


```{r}
#pc.df1 <- fread(file.dir %&%
#              "pc_projections_unrelateds_maf01_pop-info.txt")
```


## Create PC Plots color-coded by population/ancestry

### PC Analysis 1

```{r}
full.samples.hgdp <- (pc.df1 %>% 
  filter(.,(grepl(x=sample.ID,pattern="HGDP"))))$sample.ID # 929
full.samples.mcps <- (pc.df1 %>% 
  filter(.,(grepl(x=sample.ID,pattern="MCPS"))))$sample.ID # 138511
full.samples.1kg.hg <- (pc.df1 %>% filter(.,!(grepl(x=sample.ID,pattern="HGDP"))) %>% 
  filter(.,grepl(x=sample.ID,pattern="HG")))$sample.ID # 1656 
full.samples.1kg.na <- (pc.df1 %>% filter(.,!(grepl(x=sample.ID,pattern="HGDP"))) %>% 
  filter(.,grepl(x=sample.ID,pattern="NA")))$sample.ID # 848 
remove.vec.full <- c(full.samples.hgdp,full.samples.mcps,
                     full.samples.1kg.hg,full.samples.1kg.na)
full.samples.mais <- filter(pc.df1,!(sample.ID%in%remove.vec.full))$sample.ID
```


Create Plot ID variable
```{r}
plot.ID1 <- c()
plot.ID2 <- c()
pb <- txtProgressBar(min=0,max=dim(pc.df1)[1],style=3)
for (i in 1:dim(pc.df1)[1]){
  setTxtProgressBar(pb,i)
  row.df <- pc.df1[i,] %>%
    dplyr::select(one_of("sample.ID","population","region","sample.type"))
  samp.id <- row.df$sample.ID
  source <- ifelse(samp.id %in% full.samples.hgdp,"HGDP",
             ifelse(samp.id %in% full.samples.1kg.hg,"KG",
              ifelse(samp.id %in% full.samples.1kg.na,"KG",
               ifelse(samp.id %in% full.samples.mais,"MAIS","MCPS"))))
  region <- row.df$region
  if (source=="HGDP"){
    pop <- ifelse(region=="EUROPE","EUR",
           ifelse(region=="AFRICA","AFR",
           ifelse(region=="AMERICA","AMR",
           ifelse(region=="EAST_ASIA","EAS",
           ifelse(region=="CENTRAL_SOUTH_ASIA","C-SAS",
           ifelse(region=="OCEANIA","OCN","ME"))))))
  } else if (source=="KG"){
    pop <- ifelse(region=="EUROPE","EUR",
           ifelse(region=="AFRICA","AFR",
           ifelse(region=="AMERICA","AMR",
           ifelse(region=="EAST_ASIA","EAS","SAS"))))
  } else{
    pop <- "AMR"
  }
  model <- row.df$sample.type
  mod <- ifelse(model=="model","(M)","(P)")
  pid1 <- source %&% "_" %&% pop #%&% " " %&% mod
  pid2 <- source %&% "_" %&% pop %&% "_" %&%
    row.df$population # %&% " " %&% mod
  plot.ID1 <- append(plot.ID1,pid1)
  plot.ID2 <- append(plot.ID2,pid2)
}
pc.df1$plot.ID1 <- plot.ID1
pc.df1$plot.ID2 <- plot.ID2
```

```{r}
write.table(x=pc.df1,file=file.dir%&%
              "pc_projections_unrelateds_maf01_pop-info.txt",
            quote=F,sep="\t",row.names=F)
```

# Generate the PC scatter plots 

```{r}
col.vir <- viridis::viridis(20)
cvec <- brewer.pal(12,name="Paired")
pc.df1$plot.ID1 <- factor(pc.df1$plot.ID1,
 levels=as.character(unique(arrange(pc.df1,region,sample.type)$plot.ID1)))
reg.vec <- c("AMERICA","AFRICA","EUROPE")
```


```{r}
pc_density <- function(pc){
  plt <- ggplot(data=filter(pc.df1,population=="MCPS"),
                       aes_string(pc)) + 
    geom_density(aes(fill=plot.ID1),alpha=0.8) + 
    scale_fill_manual(values=c("gray25"),
                      name="Source",guide="none") 
    ##scale_fill_manual(values=c("gold1",cvec[11],
    ##                          "chartreuse","grey25",cvec[4],cvec[3],
    ##                          cvec[1],cvec[2]),
    ##                          name="Source",guide="none") 
  return(plt)
}

pc_density_horizontal <- function(pc,expand.vec=c(0.0,0.0)){
  pc.vector <- data.frame(pc.df1)[,names(pc.df1)==pc]
  plt <- pc_density(pc) +  
    theme(panel.border=element_blank(),panel.grid=element_blank(),
          axis.line=element_blank(),
          panel.background = element_blank(),
          axis.text=element_blank(),#text(color="white"),
          axis.ticks=element_blank(),
          axis.title=element_blank(),#text(color="white"),
          plot.margin = unit(c(0,6,-1,18), "mm")) + 
    scale_x_continuous(limits=c(min(pc.vector),max(pc.vector)),
                       expand=expand.vec)
}

pc_density_vertical <- function(pc,expand.vec=c(0.0,0.0)){
  pc.vector <- data.frame(pc.df1)[,names(pc.df1)==pc]
  plt <- pc_density(pc) +  
    theme(panel.border=element_blank(),panel.grid=element_blank(),
          axis.line=element_blank(),
          panel.background = element_blank(),
          axis.text=element_blank(),#text(color="white"),
          axis.ticks=element_blank(),
          axis.title=element_blank(),#text(color="white"),
          plot.margin = unit(c(7,0,14,-3), "mm")) + 
    scale_x_continuous(limits=c(min(pc.vector),max(pc.vector)),
                       expand=expand.vec) + coord_flip()
}
```


PC plotting function (v1) 
```{r}

pca_plot <- function(pcx,pcy){
  mcps.df <- filter(pc.df1,population=="MCPS")
  reference.df1 <- filter(pc.df1,
    plot.ID2 %in% c("KG_AFR_YRI","KG_EUR_IBS"))
  reference.df2 <- filter(pc.df1,
    plot.ID1 %in% c("MAIS_AMR"))
  reference.df2$plot.ID2 <- "MAIS_AMR"
  reference.df <- rbind(reference.df1,reference.df2)
  pc.plt <- ggplot(data=mcps.df,
                aes_string(x=pcx,y=pcy)) +
  geom_point(shape=21,fill="grey25",aes(alpha=sample.type),
             size=2.5) +
  geom_point(data=reference.df,
             shape=21,aes(fill=plot.ID2,alpha=sample.type),
             size=2.5) +
  scale_alpha_discrete(range=c(0.95,0.3),name="Sample Type",
    guide = guide_legend(override.aes = list(size = 3))) + 
  scale_fill_manual(values=c(col.vir[20],col.vir[10],
                             col.vir[17]),name="Source",
    guide = guide_legend(override.aes = list(size = 3))) + 
  xlab(label=pcx%&%" (" %&% filter(pve.df1,PC==pcx)$PVE %&% "%)") +
  ylab(label=pcy%&%" (" %&% filter(pve.df1,PC==pcy)$PVE %&% "%)") +
  theme_classic() 
  
  legend.obj <- get_legend(pc.plt)
  pc.plt <- pc.plt + theme(legend.position="none")
  return(list(pc.plt,legend.obj))
}

pca_plot_set_v1 <- function(pcx,pcy){
  hist.x.plt <- pc_density_horizontal(pcx)
  hist.y.plt <- pc_density_vertical(pcy)
  pc.list <- pca_plot(pcx,pcy)
  pc.plt <- pc.list[[1]]
  # grid Extra plot 
  empty <- ggplot()+geom_point(aes(1,1), colour="white")+
            theme(axis.ticks=element_blank(), 
              panel.background=element_blank(), 
              axis.text.x=element_blank(),
              axis.text.y=element_blank(),   
              axis.title.x=element_blank(),
              axis.title.y=element_blank())
  plt.set <- grid.arrange(hist.x.plt, empty, pc.plt, hist.y.plt,
              ncol=2, nrow=2,widths=c(7, 0.3), heights=c(0.6, 7))
  complete.plt <- plot_grid(plt.set,pc.list[[2]],
                            rel_widths=c(1, 0.2))

  
}
```

Create PC plots 
```{r}
plt1a <- pca_plot_set_v1("PC1","PC2")
plt1b <- pca_plot_set_v1("PC1","PC3")
plt1c <- pca_plot_set_v1("PC2","PC3")
plt1d <- pca_plot_set_v1("PC4","PC5")
plt1e <- pca_plot_set_v1("PC6","PC7")
plt1f <- pca_plot_set_v1("PC8","PC9")
plt1g <- pca_plot_set_v1("PC10","PC11")
plt1h <- pca_plot_set_v1("PC12","PC13")
plt1i <- pca_plot_set_v1("PC14","PC15")
plt1j <- pca_plot_set_v1("PC16","PC17")
plt1k <- pca_plot_set_v1("PC18","PC19")
plt1l <- pca_plot_set_v1("PC19","PC20")
plt1m <- pca_plot_set_v1("PC2","PC4")

```

Save PC plots 
```{r}
ggsave(plot=plt1a,filename=file.dir %&% "plots/pca_model1_pc1vs2_revised.png",
       height=5,width=8.5)
ggsave(plot=plt1a,filename=file.dir %&% 
 "plots/pca_model1_pc1vs2_revised.svg",
       height=5,width=8.5)

ggsave(plot=plt1b,filename=file.dir %&% "plots/pca_model1_pc1vs3_revised.png",
       height=5,width=8.5)
ggsave(plot=plt1b,filename=file.dir %&% "plots/pca_model1_pc1vs3_revised.svg",
       height=5,width=8.5)

ggsave(plot=plt1c,filename=file.dir %&% "plots/pca_model1_pc2vs3_revised.png",
       height=5,width=8.5)
ggsave(plot=plt1c,filename=file.dir %&% "plots/pca_model1_pc2vs3_revised.svg",
       height=5,width=8.5)

ggsave(plot=plt1d,filename=file.dir %&% "plots/pca_model1_pc4vs5_revised.png",
       height=5,width=8.5)
ggsave(plot=plt1d,filename=file.dir %&% "plots/pca_model1_pc4vs5_revised.svg",
       height=5,width=8.5)

ggsave(plot=plt1e,filename=file.dir %&% "plots/pca_model1_pc6vs7_revised.png",
       height=5,width=8.5)
ggsave(plot=plt1e,filename=file.dir %&% "plots/pca_model1_pc6vs7_revised.svg",
       height=5,width=8.5)

ggsave(plot=plt1f,filename=file.dir %&% "plots/pca_model1_pc8vs9_revised.png",
       height=5,width=8.5)
ggsave(plot=plt1f,filename=file.dir %&% "plots/pca_model1_pc8vs9_revised.svg",
       height=5,width=8.5)

ggsave(plot=plt1g,filename=file.dir %&% "plots/pca_model1_pc10vs11_revised.png",
       height=5,width=8.5)
ggsave(plot=plt1g,filename=file.dir %&% "plots/pca_model1_pc10vs11_revised.svg",
       height=5,width=8.5)

ggsave(plot=plt1h,filename=file.dir %&% "plots/pca_model1_pc12vs13_revised.png",
       height=5,width=8.5)
ggsave(plot=plt1h,filename=file.dir %&% "plots/pca_model1_pc12vs13_revised.svg",
       height=5,width=8.5)

ggsave(plot=plt1i,filename=file.dir %&% "plots/pca_model1_pc14vs15_revised.png",
       height=5,width=8.5)
ggsave(plot=plt1i,filename=file.dir %&% "plots/pca_model1_pc14vs15_revised.svg",
       height=5,width=8.5)

ggsave(plot=plt1j,filename=file.dir %&% "plots/pca_model1_pc16vs17_revised.png",
       height=5,width=8.5)
ggsave(plot=plt1j,filename=file.dir %&% "plots/pca_model1_pc16vs17_revised.svg",
       height=5,width=8.5)

ggsave(plot=plt1k,filename=file.dir %&% "plots/pca_model1_pc18vs19_revised.png",
       height=5,width=8.5)
ggsave(plot=plt1k,filename=file.dir %&% "plots/pca_model1_pc18vs19_revised.svg",
       height=5,width=8.5)

ggsave(plot=plt1l,filename=file.dir %&% "plots/pca_model1_pc19vs20_revised.png",
       height=5,width=8.5)
ggsave(plot=plt1l,filename=file.dir %&% "plots/pca_model1_pc19vs20_revised.svg",
       height=5,width=8.5)

ggsave(plot=plt1m,filename=file.dir %&% "plots/pca_model1_pc2vs4_revised.png",
       height=5,width=8.5)
ggsave(plot=plt1m,filename=file.dir %&% "plots/pca_model1_pc2vs4_revised.svg",
       height=5,width=8.5)
```


## PCA Plots With MAIS regions designated 

```{r}
mais.ref.df <- fread(serv.dir%&% "shared/reference_datasets/" %&%
        "mais_information/mais-population-info_NJtree-regions.txt")
```

```{r}
pc.df2 <- pc.df1#[sample(142660,size=500,replace=F),]
pb <- txtProgressBar(min=0,max=dim(pc.df2)[1],style=3)
plot.ID1 <- c() 
for (i in 1:dim(pc.df2)[1]){
  setTxtProgressBar(pb,i)
  sub.df <- pc.df2[i,]
  if (sub.df$plot.ID1=="MAIS_AMR"){
    new.name <- "MAIS_" %&% filter(mais.ref.df,
          mais.population==sub.df$population)$Region
    plot.ID1 <- append(plot.ID1,new.name)
  } else{
    plot.ID1 <- append(plot.ID1,as.character(sub.df$plot.ID1))
  }
}
```

```{r}
write.table(x=pc.df2,file=file.dir%&%
              "pc_projections_unrelateds_maf01_pop-info_mais.txt",
            quote=F,sep="\t",row.names=F)
```


```{r}
col.vir<- viridis::viridis(20)
cvec <- brewer.pal(12,name="Paired")
cvec2 <- viridis::magma(100)
updated.levels <- c("KG_AFR","HGDP_AFR","KG_EUR","HGDP_EUR",
    "KG_AMR","HGDP_AMR","MAIS_North","MAIS_Northwest",
    "MAIS_Central","MAIS_South","MAIS_Southeast","MCPS_AMR",
    "KG_SAS","HGDP_C-SAS","KG_EAS","HGDP_EAS","HGDP_ME","HGDP_OCN")
pc.df2$plot.ID1 <- plot.ID1
pc.df2$plot.ID1 <- factor(pc.df2$plot.ID1,
 levels=updated.levels)
##as.character(unique(arrange(pc.df2,region,
##                            sample.type)$plot.ID1)
reg.vec <- c("AMERICA","AFRICA","EUROPE")
```

Plotting functions 
```{r}
pc_density_v2 <- function(pc){
  ###plt <- ggplot(data=filter(pc.df2,region%in%reg.vec),
  ###                     aes_string(pc)) +
  plt <- ggplot(data=filter(pc.df2,population=="MCPS"),
                       aes_string(pc)) + 
    geom_density(aes(fill=plot.ID1),alpha=0.8) + 
    scale_fill_manual(values=c("gray25"),
                      name="Source",guide="none") 
    ###scale_fill_manual(values=c("gold1",cvec[11],cvec[2],cvec[1],
    ###                  cvec[4],cvec[3],cvec[6],cvec[8],
    ###                  cvec[7],cvec[5],cvec2[50],"gray25"),
    ###                  name="Source",guide="none") 
  return(plt)
}

pc_density_horizontal_v2 <- function(pc,expand.vec=c(0.0,0.0)){
  pc.vector <- data.frame(pc.df2)[,names(pc.df2)==pc]
  plt <- pc_density_v2(pc) +  
    theme(panel.border=element_blank(),panel.grid=element_blank(),
          axis.line=element_blank(),
          panel.background = element_blank(),
          #axis.text=element_text(color="white",size=0),
          axis.text=element_blank(),
          axis.ticks=element_blank(),
          axis.title=element_blank(),#text(color="white"),
          plot.margin = unit(c(0,6,-1,18), "mm")) + 
    scale_x_continuous(limits=c(min(pc.vector),max(pc.vector)),
                       expand=expand.vec)
}

pc_density_vertical_v2 <- function(pc,expand.vec=c(0.0,0.0)){
  pc.vector <- data.frame(pc.df2)[,names(pc.df2)==pc]
  plt <- pc_density_v2(pc) +  
    theme(panel.border=element_blank(),panel.grid=element_blank(),
          axis.line=element_blank(),
          panel.background = element_blank(),
          axis.text=element_blank(),   
          #element_text(color="white",size=0),
          axis.ticks=element_blank(),
          axis.title=element_blank(),#text(color="white"),
          plot.margin = unit(c(7,0,14,-3), "mm")) + 
    scale_x_continuous(limits=c(min(pc.vector),max(pc.vector)),
                       expand=expand.vec) +
  coord_flip()
}

```

```{r}
pca_plot_v2 <- function(pcx,pcy){
  pc.vec.x <- data.frame(pc.df2)[,names(pc.df2)==pcx]
  pc.vec.y <- data.frame(pc.df2)[,names(pc.df2)==pcy]
  mcps.df <- filter(pc.df2,population=="MCPS")
  reference.df1a <- filter(pc.df2,
    plot.ID2 %in% c("KG_AFR_YRI"))
  reference.df1a$plot.ID1 <- "KG_AFR_YRI"
  reference.df1b <- filter(pc.df2,
    plot.ID2 %in% c("KG_EUR_IBS"))
  reference.df1b$plot.ID1 <- "KG_EUR_IBS"
  reference.df2 <- filter(pc.df2,grepl(pattern="MAIS_",
                                       x=plot.ID1))
  reference.df <- rbind(reference.df1a,reference.df1b,reference.df2)  
  
  pc.plt <- ggplot(data=mcps.df,
                aes_string(x=pcx,y=pcy)) +
  geom_point(shape=21,fill="grey25",aes(alpha=sample.type),
             size=2.5) +
  geom_point(data=reference.df,
             shape=21,aes(fill=plot.ID1,alpha=sample.type),
             size=2.5) +
  scale_alpha_discrete(range=c(0.95,0.3),name="Sample Type",
    guide = guide_legend(override.aes = list(size = 3))) + 
  scale_fill_manual(values=c(cvec[6],cvec[8],cvec[7],
                             cvec[5],cvec2[50],
                             col.vir[20],col.vir[10]),
                    name="Source",
    guide = guide_legend(override.aes = list(size = 3))) + 
  xlab(label=pcx%&%" (" %&% filter(pve.df1,PC==pcx)$PVE %&% "%)") +
  ylab(label=pcy%&%" (" %&% filter(pve.df1,PC==pcy)$PVE %&% "%)") +
  theme_classic() + 
 scale_x_continuous(limits=c(min(pc.vec.x),max(pc.vec.x)))+
 scale_y_continuous(limits=c(min(pc.vec.y),max(pc.vec.y)))  
  legend.obj <- get_legend(pc.plt)
  pc.plt <- pc.plt + theme(legend.position="none")
  return(list(pc.plt,legend.obj))
}

pca_plot_set_v2 <- function(pcx,pcy,
              expand.vec.x=c(0.0,0.0),expand.vec.y=c(0.0,0.0)){
  hist.x.plt <- pc_density_horizontal_v2(pcx,expand.vec.x)
  hist.y.plt <- pc_density_vertical_v2(pcy,expand.vec.y)
  pc.list <- pca_plot_v2(pcx,pcy)
  pc.plt <- pc.list[[1]]
  # grid Extra plot 
  empty <- ggplot()+geom_point(aes(1,1), colour="white")+
            theme(axis.ticks=element_blank(), 
              panel.background=element_blank(), 
              axis.text.x=element_blank(),
              axis.text.y=element_blank(),   
              axis.title.x=element_blank(),
              axis.title.y=element_blank())
  plt.set <- grid.arrange(hist.x.plt, empty, pc.plt, hist.y.plt,
              ncol=2, nrow=2,widths=c(7, 0.3), heights=c(0.6, 7))
  complete.plt <- plot_grid(plt.set,pc.list[[2]],
                            rel_widths=c(1, 0.2))

}
```


Generate PC plots 
```{r}
plt2a <- pca_plot_set_v2("PC1","PC2")#,c(0.06,0.06),c(0.11,0.11))
plt2b <- pca_plot_set_v2("PC1","PC3")#,#c(0.1,0.0),c(0.06,0.1))
plt2c <- pca_plot_set_v2("PC1","PC4")
plt2d <- pca_plot_set_v2("PC2","PC3")
plt2e <- pca_plot_set_v2("PC2","PC4")#,c(0.1,0.1),c(0.06,0.06))
plt2f <- pca_plot_set_v2("PC3","PC4")
```

Save PC plots 
```{r}
ggsave(plot=plt2a,filename=file.dir %&% "plots/pca_model1_pc1vs2_NJregions_revised.png",
       height=5,width=8.5)
ggsave(plot=plt2a,filename=file.dir %&% 
 "plots/pca_model1_pc1vs2_NJregions_revised.svg",
       height=5,width=8.5)

ggsave(plot=plt2b,filename=file.dir %&% "plots/pca_model1_pc1vs3_NJregions_revised.png",
       height=5,width=8.5)
ggsave(plot=plt2b,filename=file.dir %&% 
 "plots/pca_model1_pc1vs3_NJregions_revised.svg",
       height=5,width=8.5)

ggsave(plot=plt2c,filename=file.dir %&% "plots/pca_model1_pc1vs4_NJregions_revised.png",
       height=5,width=8.5)
ggsave(plot=plt2c,filename=file.dir %&% 
 "plots/pca_model1_pc1vs4_NJregions_revised.svg",
       height=5,width=8.5)

ggsave(plot=plt2d,filename=file.dir %&% "plots/pca_model1_pc2vs3_NJregions_revised.png",
       height=5,width=8.5)
ggsave(plot=plt2d,filename=file.dir %&% 
 "plots/pca_model1_pc2vs3_NJregions_revised.svg",
       height=5,width=8.5)

ggsave(plot=plt2e,filename=file.dir %&% "plots/pca_model1_pc2vs4_NJregions_revised.png",
       height=5,width=8.5)
ggsave(plot=plt2e,filename=file.dir %&% 
 "plots/pca_model1_pc2vs4_NJregions_revised.svg",
       height=5,width=8.5)

ggsave(plot=plt2f,filename=file.dir %&% "plots/pca_model1_pc3vs4_NJregions_revised.png",
       height=5,width=8.5)
ggsave(plot=plt2f,filename=file.dir %&% 
 "plots/pca_model1_pc3vs4_NJregions_revised.svg",
       height=5,width=8.5)
```


